The `ordertrans()` function

### 2021-01-30 inst/ordertrans.Rmd
better default for players argument to ordertrans()
bugfix and typos

---
title: "The ordertrans() function"
output: html_document
---

@@

The `ordertrans()` function can be difficult to understand and this
short document provides some sensible use-cases.

## Hello world

The manpage provides a very simple example:

@@

```{r,label=useordertrans}
o <- ordertrans(x,letters[1:4])
o
```

@@

Let us consider the skating dataset and use `ordertrans()` to study
it.

@@

We might ask how judges 1 and 2 compare to one another?  We need to
create vectors like `x` and `y` above:

@@

In the above, see how objects `j1` and `j2` have identical names, in
the same order.  Observe that `hughes` is ranked 1 by judge 1, and 4
by judge 2.  This makes it sensible to plot `j1` against `j2`:

```{r}
par(pty='s')  # forces plot to be square
plot(j1,j2,asp=1,pty='s',xlim=c(0,25),ylim=c(0,25),pch=16,xlab='judge 1',ylab='judge 2')
abline(0,1)  # diagonal line
for(i in seq_along(j1)){text(j1[i],j2[i],names(j1)[i],pos=4,col='gray',cex=0.7)}
```

In the above plot we see general agreement but differences in detail.
For example, `hughes` is ranked first by judge 1 and fourth by judge
2.  However, other problems are not so easy.  Suppose we wish to
compare the ranks according to likelihood with the ranks according to
some points system.

@@

```{r,label=crapplot}
plot(mL,ordertrans(mP,mL))
```

However, the plot above is a bit crude.  Function `ordertransplot()`
gives a more visually pleasing output:

```{r,label=showoffordertransplot}
ordertransplot(mL,mP,xlab="likelihood rank",ylab="Borda rank")
```

@@

```{r,transplotjudge1}
ordertransplot(mL,j1,xlab="likelihood rank",ylab="Judge 1 rank")
```

In the above, looking at the lower-left corner, we see (reading
horizontally) that the likelihood method placed Slutskya first, then
Hughes second, then Kwan third; while (reading vertically) judge 1
placed Hughes first, then Kwan, then Slutskya.

>>>>

---
title: "The `ordertrans()` function"
output: bookdown::html_document2
---

@@

The `ordertrans()` function can be difficult to understand and this
short document provides some sensible use-cases.  The manpage provides
a very simple example, but here we are going to use an even simpler
example:

@@

```{r,label=useordertrans}
o <- ordertrans(x) # by default, sorts names() into alphabetical order
o
```

@@

Let us consider the skating dataset and use `ordertrans()` to study it
(note that the skating dataset is analysed in more depth in
`skating.Rmd`).

@@

We might ask how judges `J1` and `J2` compare to one another?  We need to
create vectors like `x` and `y` above:

@@

In the above, see how objects `j1` and `j2` have identical names, in
the same order.  Observe that `hughes` is ranked 1 (that is, first) by
`J1`, and 4 (that is, 4th) by `J2`.  This makes it sensible to plot
`j1` against `j2`:

```{r j1vsj2,fig.cap="Judge 1 vs judge 2"}
par(pty='s')  # forces plot to be square
plot(j1,j2,asp=1,pty='s',xlim=c(0,25),ylim=c(0,25),pch=16,xlab='judge 1',ylab='judge 2')
abline(0,1)  # diagonal line
for(i in seq_along(j1)){text(j1[i],j2[i],names(j1)[i],pos=4,col='gray',cex=0.7)}
```

In figure \@ref(fig:j1vsj2), we see general agreement but differences
in detail.  For example, `hughes` is ranked first by judge 1 and
fourth by judge 2.  However, other problems are not so easy.  Suppose
we wish to compare the ranks according to likelihood with the ranks
according to some points system.

@@

```{r,label=crapplot,fig.cap="points-based ranks vs likelihood ranks"}
plot(mL,ordertrans(mP,names(mL)))
```

However, figure \@ref(fig:crapplot) is a bit crude.  Function
`ordertransplot()` gives a more visually pleasing output, see figure
\@ref(fig:showoffordertransplot).

```{r,label=showoffordertransplot,fig.cap="points=based rank vs likelihood rank using `ordertransplot()`"}
ordertransplot(mL,mP,xlab="likelihood rank",ylab="Borda rank")
```

@@

```{r transplotjudge1,fig.cap="Likelihood rank vs rank according to Judge 1"}
ordertransplot(mL,j1,xlab="likelihood rank",ylab="Judge 1 rank")
```

In figure \@ref(fig:transplotjudge1), looking at the lower-left
corner, we see (reading horizontally) that the likelihood method
placed Slutskya first, then Hughes second, then Kwan third; while
(reading vertically) judge 1 placed Hughes first, then Kwan, then
Slutskya.
